\vspace{-.2in} Squint: Lossy Hierarchical Compression on Symbol Streams \vspace{-.1in}

### 2012-12-12 tex/paper.tex
Some formatting

Given an input of length $n$, let $r$ be the number of rules generated by
\Sequitur from the input.  In the analysis of the \Similarity algorithm, let $R$
be the total number of variable replacements we perform. For each replacement,
we had to have searched though up to all pairs of variables, and for each pair
of variables expanding them and calculating the levenshtein distance could take
time linear in $n$.  Performing the replacement only takes time linear in $r$.
Thus the runtime is $\boxed{O(R*r^2*n)}$.

From the \Sequitur paper \cite{sequitur} we know that the number of rules in
the grammar generated by \Sequitur is bounded by $O(n)$. Thus the worst
case runtime on the similarity algorithm is $\boxed{O(n^4)}$. This is because
we could have up to $r$ replacements.  However, in practice we will see that
the runtime is much better than $O(n^4)$. There are a variety of possible
reasons for this: in our test data, $r$ didn't grow quite linearly in $n$, $R$
could grow much slower than $r$, and the use of native code in calculating the
levenshtein distance made the its contribution much less noticeable for the
input sizes we tested.

@@

In the tables, Time is the
wall clock running as measured on a 2.26 Ghz Intel Core 2 Duo Macbook Pro.
Ratio is the relative size of the grammar produced compared with \Sequitur,
where the size of the grammar is the sum of the lengths of the rhs's of the
rules. Levdistance is the levenshtein distance between the output of the
lossified grammar and the original input string, it is a kind of measure of the
fidelity of our algorithms but doesn't take into account how well patterns were
preserved. For the rep-n tests, proportion is the proportion of the final
string produced that was `*'. The closer to 1/30~0.033, the more accurate the
lossifier was in recreating the original texture of the input.

@@

Across all of the tests, the leveshtein
distances were comparable for both algorithms. Though they differed in
individual cases, no algorithm did consistently better than the other.
In general the levenshtein distances were high, except in the $\epsilon=.2$
case for the gen-n data where very little compression was possible at all.
We believe most of the distance could be explained by
the fact that the algorithms could drastically shorten
the strings, since it always replaces variables with \emph{smaller} variables
as discussed earlier.
The proportion of '*'s in the rep-n input
was usually the right order of magnitude, with both algorithms making big
mistakes occassionally. In most cases, we felt that the two algorithms maintained similarly
acceptable fidelity levels in recreating the texture of the original rep-n
input file, and we discuss this subjective impression further in the
Larger Tests section.

@@

In summary, both of the algorithms were able to achieve moderate additional compression ratios
on top of \Sequitur. The fidelity between the original and lossified strings
wasn't great as measured by levenshtein distance, but we think the texture in both the
rep-n and gen-n datasets was preserved well. Both fidelity and compression ratio were
very sensitive to $\epsilon$ depending on the kind of data.
The differences between the two algorithms lie almost entirely in their
runtime. Empirically the \Similarity algorithm tends to
consistently incur an overhead of about $O(\sqrt{n})$ on top of the runtime for
the \Cluster algorithm. Because of this we dismiss the \Similarity algorithm as
unscalable and do not test it further.

@@

\subsection*{deckmyn.c}

@@

\subsection*{gen256000 Output String, Cluster, $\epsilon=.4$:}

>>>>

Given an input of length $n$, let $r$ be the number of rules generated by
\Sequitur from the input.  In the analysis of the \Similarity algorithm, let $R$
be the total number of variable replacements we perform. For each replacement,
we had to have searched though up to all pairs of variables, and for each pair
of variables expanding them and calculating the Levenshtein distance could take
time linear in $n$.  Performing the replacement only takes time linear in $r$.
Thus the runtime is $\boxed{O(R*r^2*n)}$.

From the \Sequitur paper \cite{sequitur} we know that the number of rules in
the grammar generated by \Sequitur is bounded by $O(n)$. Thus the worst
case runtime on the similarity algorithm is $\boxed{O(n^4)}$. This is because
we could have up to $r$ replacements.  However, in practice we will see that
the runtime is much better than $O(n^4)$. There are a variety of possible
reasons for this: in our test data, $r$ didn't grow quite linearly in $n$, $R$
could grow much slower than $r$, and the use of native code in calculating the
Levenshtein distance made the its contribution much less noticeable for the
input sizes we tested.

@@

In the tables, Time is the wall clock running as measured on a 2.26 Ghz Intel
Core 2 Duo Macbook Pro.  Ratio is the relative size of the grammar produced
compared with \Sequitur, where the size of the grammar is the sum of the
lengths of the rhs's of the rules. Levdistance is the Levenshtein distance
between the output of the lossified grammar and the original input string, it
is a kind of measure of the fidelity of our algorithms but doesn't take into
account how well patterns were preserved. For the rep-n tests, proportion is
the proportion of the final string produced that was `*'. The closer to
1/30~0.033, the more accurate the lossifier was in recreating the original
texture of the input.

@@

Across all of the tests, the Leveshtein distances were comparable for both
algorithms. Though they differed in individual cases, no algorithm did
consistently better than the other.  In general the Levenshtein distances were
high, except in the $\epsilon=.2$ case for the gen-n data where very little
compression was possible at all.  We believe most of the distance could be
explained by the fact that the algorithms could drastically shorten the
strings, since it always replaces variables with \emph{smaller} variables as
discussed earlier.  The proportion of '*'s in the rep-n input was usually the
right order of magnitude, with both algorithms making big mistakes
occassionally. In most cases, we felt that the two algorithms maintained
similarly acceptable fidelity levels in recreating the texture of the original
rep-n input file, and we discuss this subjective impression further in the
Larger Tests section.

@@

In summary, both of the algorithms were able to achieve moderate additional
compression ratios on top of \Sequitur. The fidelity between the original and
lossified strings wasn't great as measured by Levenshtein distance, but we
think the texture in both the rep-n and gen-n datasets was preserved well. Both
fidelity and compression ratio were very sensitive to $\epsilon$ depending on
the kind of data.  The differences between the two algorithms lie almost
entirely in their runtime. Empirically the \Similarity algorithm tends to
consistently incur an overhead of about $O(\sqrt{n})$ on top of the runtime for
the \Cluster algorithm. Because of this we dismiss the \Similarity algorithm as
unscalable and do not test it further.

@@

\section{Large Examples}

\subsection*{deckmyn.c, Cluster, $\epsilon=.4$:}

@@

\vspace{0.2in}
\subsection*{gen256000, Cluster, $\epsilon=.4$:}
